---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
library(lubridate)
activity$processedday <- wday(ymd(activity$date),label=TRUE)

```



## What is mean total number of steps taken per day?

```{r}
StepsPerDay <- tapply(activity$steps,activity$date,sum,na.rm=TRUE)
hist(StepsPerDay,main="Histogram of Steps Per Day",xlab="Steps Per Day")
```

The mean total number of steps taken per day is `r mean(StepsPerDay,na.rm=TRUE)` and the median is `r median(StepsPerDay,na.rm=TRUE) `

## What is the average daily activity pattern?
```{r}
StepsPerInterval <-tapply(activity$steps,activity$interval,mean,na.rm=TRUE)
plot(StepsPerInterval,type="l",main="Steps Per Interval",xlab="Interval",ylab="Steps")
```


The 5 minute interval with the maximum number of steps on average is `r which.max(StepsPerInterval)`


## Imputing missing values

Here we impute missing values by filling with average for the interval the missing value is in.

```{r}
activity[is.na(activity$steps),'steps'] <- sapply(activity[is.na(activity$steps),'interval'],function (x) { StepsPerInterval[as.character(x)]})
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
par(mfrow=c(2,1))

weekendactivity <- activity[activity$processedday=='Sat' | activity$processedday=='Sun',]
weekdayactivity <- activity[activity$processedday!='Sat' & activity$processedday!='Sun',]
weekendStepsPerInterval <- tapply(weekendactivity$steps,weekendactivity$interval,mean)
weekdayStepsPerInterval <- tapply(weekdayactivity$steps,weekdayactivity$interval,mean)
plot(weekendStepsPerInterval,type='l',main="Weekend Steps Per Interval",xlab="Interval", ylab="Steps")
plot(weekdayStepsPerInterval,type='l',main="Weekday Steps Per Interval",ylab="Steps",xlab="Interval")


```
